\Title

### 2023-06-01 Grimaldi-etal-BiolCybernetics.tex
Results : C

We observe some interesting facts. First, as we change the mean spatial frequency of the texture, we observe a monotonic decrease in accuracy. This comes as a similar trend as shown in the primary visual areas~\citep{priebe_tuning_2006} and reveals the most informative scales learned by our model. We observe the trend for different spatial frequency bandwidths. In particular, the accuracy is better for a grating-like stimulus than for a large one (which qualitatively resembles a more textured stimulus), reminiscent of the behavioral response of humans to such stimuli~\citep{simoncini_more_2012, ravello_speed-selectivity_2019}. Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth. When the stimulus is grating-like and the orientation is arbitrary with respect to the direction of motion, the system faces the aperture problem and experiences a decrease in accuracy. This is not the case for isotropic stimuli or when the orientation is perpendicular to the direction of motion. Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter. This shows a progressive decrease in accuracy, similar to that observed in the amplitude of human eye movements~\citep{mansour_pour_speed_2018}.
%
\subsection{Accuracy efficiency trade-off}% for the motion detection task}
%%%-----------------------------------------------------------------
Once our MLR model is trained, we obtain spatiotemporal kernels corresponding to the weights associated with the heterogeneous delays of our layer of spiking neurons, which can be used for detection. We observed that the distribution of the kernel weights is sparse, with most values close to zero (\seeFig{kernels}). As shown in the formalization of our event-based model, the computational cost of our model, if implemented on a neuromorphic chip, would be dominated by the computations used for the convolution operation. In a dense setting, this corresponds for all voxels in the output to a sum over all voxels in the inputs for all weights in the kernel. If the information support is sparse, then computations can only be performed on those events. Also, if we set some weights of the kernels to zero, then the sum can be skipped for those addresses. Thus, knowing the sparseness of the input, the total number of computations scales with the number of spikes multiplied by the number of non-zero synaptic weights.

To assess the robustness of the classification as a function of computational load, we will prune the weights in $\{\synapse_\ranksyn\}_{\ranksyn \in [0,\Nsyn)}$ that are below a defined threshold. In Fig.~\ref{fig:accuracy} we plot the classification accuracy as a function of the relative number of computations or active weights per decision for each neuron in the layer. For comparison, and to take into account the performance gain from using heterogeneous delays, we show the accuracy obtained with an MLR model using a 2D time surface (in red) as in~\citep{grimaldi_robust_2022}. This latter method is based on the delays of the last recorded events and requires fewer computations (in our case $17\times17$) than the dense 3D kernels without any pruning ($21\times17\times17$). While fewer computations are needed, the classification performance obtained for the model using time surfaces is similar to our method using all the weights of the kernels.

By pruning the weights, we observe that the evolution of accuracy as a function of the log percentage of active weights fits well to a sigmoid curve. Half saturation is reached at a ratio of about $3.2\ times 10^{-2}$ of active weights, which in our setting corresponds to a total amount of about $180$ computations per decision. Compared to the full kernels, the accuracy of our method is maintained to its top performance when the number of computations is divided by a factor up to about $31$. In this case, the number of computations is greatly reduced compared to~\citep{grimaldi_robust_2022}, thus demonstrating the efficiency of the presented method.
%
%%%-----------------------------------------------------------------
%: fig:accuracy
%%%-----------------------------------------------------------------
\begin{figure}%[h!]
    \centering
    \includegraphics[width=0.95\linewidth]{figures/quant_accuracy.pdf}
    \caption{
        Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line). The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel. As we prune the coefficients, we observe a stable accuracy value, with a drop observed at about 25 times fewer computations.
        }
    \label{fig:accuracy}
\end{figure}
%
\section{Discussion}
%%%-----------------------------------------------------------------
In this paper, we have introduced a generic SNN with heterogeneous delays and shown how it compares favorably to a state-of-the-art event-based classification algorithm for a visual motion detection task. The learned model shares a number of similarities with neurobiological anatomical observations, as well as with behavioral results. The event-based computations of our method can be drastically reduced by pruning synapses, while maintaining top classification performance. This shows that we can take advantage of the precise timing of spikes to improve the performance of neural computations.
%
\subsection{Synthesis and main contributions}
%%%-----------------------------------------------------------------
The HD-SNN model that has been trained and evaluated on a naturalistic motion detection task. The model has been defined to provide optimal detection of event-driven spatiotemporal motifs. We have shown that the model, when trained on a dataset of natural images with realistic eye movements, learns kernels similar to those found in the early visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}. %We then demonstrated that the model shows similarities to biological responses.
We have evaluated the computational cost of this model when implemented in a setting similar to event-based hardware. We show that the use of heterogeneous delays may be an efficient computational solution for future neuromorphic hardware, but also a key to understanding why spikes are a universal component of neural information processing.

>>>>

We observe some interesting facts. First, as we change the mean spatial frequency of the texture, we observe a monotonic decrease in accuracy. This comes as a similar trend as shown in the primary visual areas~\citep{priebe_tuning_2006} and reveals the most informative scales learned by our model. We observe the trend for different spatial frequency bandwidths. In particular, the accuracy is better for a grating-like stimulus than for a large one (which qualitatively resembles a more textured stimulus), reminiscent of the behavioral response of humans to such stimuli~\citep{simoncini_more_2012, ravello_speed-selectivity_2019}. Interestingly, we also see a modulation of accuracy as a function of orientation bandwidth. When the stimulus is grating-like and the orientation is arbitrary with respect to the direction of motion, the system faces the aperture problem and experiences a decrease in accuracy. This is not the case for isotropic stimuli or when the orientation is perpendicular to the direction of motion. Finally, we manipulated the amount of change between two successive frames, similar to a temperature parameter. This shows a progressive decrease in accuracy, similar to that observed in the amplitude of human eye movements~\citep{mansour_pour_speed_2018}.
%
\subsection{Accuracy efficiency trade-off}% for the motion detection task}
%%%-----------------------------------------------------------------
After training our MLR model, we obtain spatiotemporal kernels corresponding to the weights associated with the heterogeneous delays of our layer of spiking neurons, which can be used for detection. We observed that the distribution of the kernel weights is sparse, with most values close to zero (\seeFig{kernels}). As shown in the formalization of our event-based model, the computational cost of our model, if implemented on a neuromorphic chip, would be dominated by the computations used for the convolution operation. In a dense setting, this corresponds for all voxels in the output to a sum over all voxels in the inputs for all weights in the kernel. If the information support is sparse, then computations can only be performed on those events. Also, if we set some weights of the kernels to zero, then the sum for those addresses can be dropped. Thus, knowing the sparseness of the input, the total number of computations scales with the number of spikes multiplied by the number of non-zero synaptic weights.

To assess the robustness of the classification as a function of computational load, we will prune the weights in $\{\synapse_\ranksyn\}_{\ranksyn \in [0,\Nsyn)}$ that are below a defined threshold. In Fig.~\ref{fig:accuracy} we plot the classification accuracy as a function of the relative number of computations or active weights per decision for each neuron in the layer. For comparison, and to take into account the performance gain from using heterogeneous delays, we show the accuracy obtained with an MLR model using a 2D time surface (in red) as in~\citep{grimaldi_robust_2022}. The latter method is based on the delays of the last recorded events and requires fewer computations (in our case $17\times17$) than the dense 3D kernels without any pruning ($21\times17\times17$). While fewer computations are needed, the classification performance obtained for the model using time surfaces is similar to our method using all the weights of the kernels.

By pruning the weights, we observe that the evolution of accuracy as a function of the log percentage of active weights fits well to a sigmoid curve. Half saturation is reached at a ratio of about $3.2\ times 10^{-2}$ of active weights, which in our setting corresponds to a total of about $180$ computations per decision. Compared to the full kernels, the accuracy of our method is maintained relative to its peak performance when the number of computations is divided by a factor up to about $31$. In this case, the number of computations is greatly reduced compared to~\citep{grimaldi_robust_2022}, demonstrating the efficiency of the presented method.
%
%%%-----------------------------------------------------------------
%: fig:accuracy
%%%-----------------------------------------------------------------
\begin{figure}%[h!]
    \centering
    \includegraphics[width=0.95\linewidth]{figures/quant_accuracy.pdf}
    \caption{
        Accuracy as a function of computational load for the HD-SNN model (blue dots) with error bars indicating the 5\% - 95\% quantiles and a sigmoid fit (blue line). The relative computational load (on a logarithmic axis) is controlled by changing the percentage of nonzero weights relative to the dense convolution kernel. As we prune the coefficients, we observe a stable accuracy value, with a drop observed at about 25 times fewer computations.
        }
    \label{fig:accuracy}
\end{figure}
%
\section{Discussion}
%%%-----------------------------------------------------------------
In this paper, we have introduced a generic SNN with heterogeneous delays and shown how it compares favorably to a state-of-the-art event-based classification algorithm for a visual motion detection task. The learned model shares a number of similarities with neurobiological anatomical observations, as well as with behavioral results. The event-based computations of our method can be drastically reduced by pruning synapses, while maintaining top classification performance. This shows that we can take advantage of the precise timing of spikes to improve the performance of neural computations.
%
\subsection{Synthesis and main contributions}
%%%-----------------------------------------------------------------
The HD-SNN model that has been trained and evaluated on a naturalistic motion detection task. The model has been defined to provide optimal detection of event-driven spatiotemporal motifs. We have shown that the model, when trained on a dataset of natural images with realistic eye movements, learns kernels similar to those found in the early visual cortex~\cite{deangelis_functional_1999, kremkow_push-pull_2016}. %We then demonstrated that the model shows similarities to biological responses.
We have evaluated the computational cost of this model when implemented in a setting similar to event-based hardware. We show that the use of heterogeneous delays may be an efficient computational solution for future neuromorphic hardware, but also a key to understanding why spikes are a universal component of neural information processing.
